Tail Indices and Scale Parameters in Financial Time Series

### 2017-03-22 papers/FX/draft.tex
Revised the scripts to add error bars.

\subsection{Hybrid fitting of the shifted Pareto distribution}
\label{sec:hybrid_estimation}
We adopt a hybrid procedure to estimate the parameters of the
distribution function assumed in \eqref{eq:X_distr}: First we
obtain a Hill estimate of the lower-tail indices of an
equity return and fix $\alpha$ in \eqref{eq:X_distr}
to its Hill estimate. Then we estimate $K$ by maximizing
the likelihood function with fixed $\alpha$. Obviously we take only
negative observations and hence condition on $X < 0$.

@@

\subsection{Pickand's Estimator}
Pickands (1975) \cite{pickands1975statistical} proposed a different
approach to estimate the tail index than that of Hill. Assume
$\{x_i\}_{i=1}^n$ are samples from the distribution of $X$ and let $U(t)$
denote the quantile of $1 - 1/t$ of this distribution, i.e.
\[
P(X > U(t)) = {1 \over t}
\]
The key idea of Pickands' estimator is to observe
\[
\lim_{t \to \infty} {
  U(c(t) t) - U(t)
  \over
  U(t) - U(t/c(t))
} = 2^{1/\alpha}
\]
for any function $c(t)$ satisfying $c(t) \to 2$ as $t \to \infty$.
Pickands formulated the estimation problem in terms of the extreme
value index $\xi$, which is the reciprocal of the tail index $\alpha$
when the distribution of $X$ is regularly varying (cf. Embrechtsm,
Kl\"uppelberg and Mikosch \cite{Embrechts1997}).

Building upon the aforementioned observation and taking $c(t)$ to
be simply constant 2, Pickands proposed the following estimator for
$\xi$:
\[
\hat \xi = {1 \over \ln 2} \ln {
  x_{(k)} - x_{(2k)}
  \over
  x_{(2k)} - x_{(4k)}
}
\]
Dekkers and de Haan (1989) later showed \cite{dekkers1989estimation}
that Pickands' estimator was consistent in both the weak and the strong
sense, when the sequence $k(n)$ was chosen appropriately:
\begin{itemize}
\item if $k \to \infty$ and $k/n \to 0$ as $n \to \infty$,
  \[
  \hat \xi \overset{P}{\to} \xi
  \]
\item if $k \to \infty$, $k/n \to 0$ and $k / \ln(\ln n) \to \infty$
  as $n \to \infty$,
  \[
  \hat \xi \overset{a.s.}{\to} \xi
  \]
\end{itemize}
Dekkers and de Haan \cite{dekkers1989estimation} also showed the
asymptotic normality of the estimator, when some
extra technical conditions were satisfied by
$U(\cdot)$ and $k(n) \to \infty$ yet grew sufficiently slowly, i.e.
\[
\sqrt k (\hat \xi - \xi) \overset{d}{\to} N(0, v(\xi))
\]
where
\[
v(\xi) = {
  \xi^2 (2^{2 \xi + 1} + 1)
  \over
  [2 \ln(2) (2^\xi - 1)]^2
}
\]
Using Pickands' estimator, the same result of very similar tail indices
is again obtained for the {\it Energy} and the {\it Consumer Staples}
sectors. Figure \ref{fig:Energy_Pickands} and
\ref{fig:Consumer_Staples_Pickands} illustrate this claim.
\begin{figure}[htb!]
  \includegraphics[width=\textwidth]{Energy_Pickands.pdf}
  \caption{Pickands' estimates of {\it Energy} stocks.}
  \label{fig:Energy_Pickands}
\end{figure}

\begin{figure}[htb!]
  \includegraphics[width=\textwidth]{Consumer_Staples_Pickands.pdf}
  \caption{Pickands' estimates of {\it Consumer Staples} stocks.}
  \label{fig:Consumer_Staples_Pickands}
\end{figure}

>>>>

\subsection{Hybrid fitting of the shifted Pareto distribution}
\label{sec:hybrid_estimation}
We adopt a hybrid procedure to estimate the parameters of the
distribution function assumed in \eqref{eq:X_distr}: First we
obtain a Hill estimate of the lower-tail indices of an
equity return and fix $\alpha$ in \eqref{eq:X_distr}
to its Hill estimate. Then we estimate $K$ by maximizing
the likelihood function with fixed $\alpha$. Obviously we take only
negative observations and hence condition on $X < 0$. Regularity
conditions on the maximum likelihood estimator are easily checked for
the distribution \eqref{eq:X_distr}.
The density function of $X$ conditional on $X < 0$ is
\[
f_L(x) = {\alpha K^{\alpha} \over (K - x)^{\alpha + 1}}
\]
Clearly, the log-likelihood function of the $K$ parameter is
\[
l(K | X < 0) = \ln(\alpha) + \alpha \ln(K) - (\alpha + 1)\ln(K - X)
\]
where, as is just mentioned, we pre-estimate $\alpha$ with Hill
estimator and regard it as known at this step. With $n$ observations
$X_1, \dots, X_n$, the log-likelihood function of $K$ is
\[
L(K | \forall i=1,2,...,n, X_i < 0) = n \ln(\alpha) + n \alpha \ln(K)
- (\alpha + 1) \sum_{i=1}^n \ln(K - X_i)
\]
To obtain the maximum likelihood estimation, we need to solve the
equation $\pd{L}{K} = 0$, which expands to
\[
{n \alpha \over K} - (\alpha + 1) \sum_{i=1}^n {1 \over K - X_i} = 0
\]
This equation has to be solved numerically.
The Fisher information of $K$ follows as
\begin{eqnarray*}
  I(K) &=& \E\left[ \left( \opd{K} l(K | X < 0) \right)^2 \right] \\
  &=& \alpha K^{\alpha - 2}
  \int_{-\infty}^0 {(\alpha x + K)^2 \over (K - x)^{\alpha + 3}} dx \\
  &=& {\alpha \over (2 + \alpha) K^2}
\end{eqnarray*}
So the maximum likelihood estimator, call it $\hat K$, is
asymptotically normal:
\[
\sqrt n (\hat K - K) \to N \left(
  0, \left( 1 + {2 \over \alpha} \right) K^2
\right)
\]

@@

% \subsection{Pickand's Estimator}
% Pickands (1975) \cite{pickands1975statistical} proposed a different
% approach to estimate the tail index than that of Hill. Assume
% $\{x_i\}_{i=1}^n$ are samples from the distribution of $X$ and let $U(t)$
% denote the quantile of $1 - 1/t$ of this distribution, i.e.
% \[
% P(X > U(t)) = {1 \over t}
% \]
% The key idea of Pickands' estimator is to observe
% \[
% \lim_{t \to \infty} {
%   U(c(t) t) - U(t)
%   \over
%   U(t) - U(t/c(t))
% } = 2^{1/\alpha}
% \]
% for any function $c(t)$ satisfying $c(t) \to 2$ as $t \to \infty$.
% Pickands formulated the estimation problem in terms of the extreme
% value index $\xi$, which is the reciprocal of the tail index $\alpha$
% when the distribution of $X$ is regularly varying (cf. Embrechtsm,
% Kl\"uppelberg and Mikosch \cite{Embrechts1997}).

% Building upon the aforementioned observation and taking $c(t)$ to
% be simply constant 2, Pickands proposed the following estimator for
% $\xi$:
% \[
% \hat \xi = {1 \over \ln 2} \ln {
%   x_{(k)} - x_{(2k)}
%   \over
%   x_{(2k)} - x_{(4k)}
% }
% \]
% Dekkers and de Haan (1989) later showed \cite{dekkers1989estimation}
% that Pickands' estimator was consistent in both the weak and the strong
% sense, when the sequence $k(n)$ was chosen appropriately:
% \begin{itemize}
% \item if $k \to \infty$ and $k/n \to 0$ as $n \to \infty$,
%   \[
%   \hat \xi \overset{P}{\to} \xi
%   \]
% \item if $k \to \infty$, $k/n \to 0$ and $k / \ln(\ln n) \to \infty$
%   as $n \to \infty$,
%   \[
%   \hat \xi \overset{a.s.}{\to} \xi
%   \]
% \end{itemize}
% Dekkers and de Haan \cite{dekkers1989estimation} also showed the
% asymptotic normality of the estimator, when some
% extra technical conditions were satisfied by
% $U(\cdot)$ and $k(n) \to \infty$ yet grew sufficiently slowly, i.e.
% \[
% \sqrt k (\hat \xi - \xi) \overset{d}{\to} N(0, v(\xi))
% \]
% where
% \[
% v(\xi) = {
%   \xi^2 (2^{2 \xi + 1} + 1)
%   \over
%   [2 \ln(2) (2^\xi - 1)]^2
% }
% \]
% Using Pickands' estimator, the same result of very similar tail indices
% is again obtained for the {\it Energy} and the {\it Consumer Staples}
% sectors. Figure \ref{fig:Energy_Pickands} and
% \ref{fig:Consumer_Staples_Pickands} illustrate this claim.
% \begin{figure}[htb!]
%   \includegraphics[width=\textwidth]{Energy_Pickands.pdf}
%   \caption{Pickands' estimates of {\it Energy} stocks.}
%   \label{fig:Energy_Pickands}
% \end{figure}

% \begin{figure}[htb!]
%   \includegraphics[width=\textwidth]{Consumer_Staples_Pickands.pdf}
%   \caption{Pickands' estimates of {\it Consumer Staples} stocks.}
%   \label{fig:Consumer_Staples_Pickands}
% \end{figure}
